Multiple Sclerosis Diagnosis from Natural Language Processing of Patient Text Data

### 2022-09-14 paper.tex
Added examples of diagnosis patterns

\indent The first control group constitutes random users from Twitter and Reddit who have not mentioned the term “MS” or any related terms in their timeline. The dataset was constructed such that there is an equal probability for a potential control user to be selected from Twitter or from Reddit. Moreover, we ignored image-only posts, video-only posts, Twitter retweets and Reddit cross-posts. For Reddit, we added the additional restriction that the potential control group user has never posted on any MS-related subreddit \footnote{\url{https://www.doc.ic.ac.uk/~dj321/msnlp/mssubreddits.txt}}. After identifying the users, we select random posts in their timeline to be included in our dataset. The number of posts we select per user is controlled such that the final user-comment distribution is kept the same as the diagnosed users in the dataset, as illustrated in Figure \ref{fig:usercomment}. Finally, as a measure to prevent posts created by bots to be included in our dataset, no more than 5 duplicate posts were allowed by the same user. \\

@@

\indent Therefore, the second control group are comments of users who have given self-reported diagnoses of diabetes on social media (both type 1 and type 2), and who, like the first control group, have not mentioned “MS” or any related terms in their timeline. Diabetes was chosen as the illness for the second control group because firstly, it is an incurable, chronic condition like MS, and secondly, it is relatively prevalent, so we can find enough users on social media to exactly match the number of MS-diagnosed users from the MS Register dataset. The users were found by searching on Twitter and Reddit for high-precision diagnostic patterns \footnote{List of diagnosis patterns (\url{https://www.doc.ic.ac.uk/~dj321/msnlp/diagnosispatterns.txt}) was inspired by the work of \citet{Cohan:18} on NLP-based diagnosis of mental health disorders}. After identifying these users, we only include posts where the user mentions their condition (that is, diabetes), because we would expect only those experiences to be presented to a primary healthcare provider. Then, like what we did for the diagnosed users, we remove all clauses with the word “diabetes” from the text, since the classifier will be used pre-diagnosis, and we cannot make the classifier rely on the term “diabetes” for classifying patients who do not have MS. \\

>>>>

\indent The first control group constitutes random users from Twitter and Reddit who have not mentioned the term “MS” or any related terms in their timeline. The dataset was constructed such that there is an equal probability for a potential control user to be selected from Twitter or from Reddit. Moreover, we ignored image-only posts, video-only posts, Twitter retweets and Reddit cross-posts. For Reddit, we added the additional restriction that the potential control group user has never posted on any MS-related subreddit\footnote{\url{https://www.doc.ic.ac.uk/~dj321/msnlp/mssubreddits.txt}}. After identifying the users, we select random posts in their timeline to be included in our dataset. The number of posts we select per user is controlled such that the final user-comment distribution is kept the same as the diagnosed users in the dataset, as illustrated in Figure \ref{fig:usercomment}. Finally, as a measure to prevent posts created by bots to be included in our dataset, no more than 5 duplicate posts were allowed by the same user. \\

@@

\indent Therefore, the second control group are comments of users who have given self-reported diagnoses of diabetes on social media (both type 1 and type 2), and who, like the first control group, have not mentioned “MS” or any related terms in their timeline. Diabetes was chosen as the illness for the second control group because firstly, it is an incurable, chronic condition like MS, and secondly, it is relatively prevalent, so we can find enough users on social media to exactly match the number of MS-diagnosed users from the MS Register dataset. The users were found by searching on Twitter and Reddit for high-precision diagnostic patterns\footnote{List of diagnosis patterns (\url{https://www.doc.ic.ac.uk/~dj321/msnlp/diagnosispatterns.txt}) was inspired by the work of \citet{Cohan:18} on NLP-based diagnosis of mental health disorders}. Some examples of these patterns include \emph{``i have been diagosed with diabetes"}, \emph{``he diagnosed me with diabetes"} and \emph{``i have a diagnosis of diabetes"}. After identifying these users, we only include posts where the user mentions their condition (that is, diabetes), because we would expect only those experiences to be presented to a primary healthcare provider. Then, like what we did for the diagnosed users, we remove all clauses with the word “diabetes” from the text, since the classifier will be used pre-diagnosis, and we cannot make the classifier rely on the term “diabetes” for classifying patients who do not have MS. \\
